Value Ordering, Discrepancies, and Restarts for Subgraph Algorithms\thanks{This footnote leaves room for a piece of text saying how grateful we are to our funding agency. This includes a magic grant number, so it takes up three precious lines of space.}

### 2018-01-31 paper/paper.tex
James's comments

The subgraph isomorphism problem is to find a copy of a small ``pattern'' graph inside a larger
``target'' graph.  The current single strongest subgraph isomorphism solver uses ``highest degree
first'' as a value-ordering heuristic to direct a constraint programming style search
\citep{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16,DBLP:conf/ijcai/McCreeshPT16}. This
heuristic is much better than branching randomly, but is still far from perfect.  We could try to
introduce tiebreaking mechanisms and similar refinements to strengthen the heuristic. However, given
the extreme cost of making an incorrect branching choice early on in search, we believe it is more
fruitful to investigate alternatives to simple backtracking (depth-first) search. This paper
evaluates two such approaches: discrepancy search (which performs poorly), and restarts (which are
extremely beneficial). To explore different parts of the search space when restarting, we introduce
a new value-ordering heuristic, which adds a small amount of randomness to the ``highest degree
first'' heuristic.  The effects of these changes are shown in
\cref{figure:scatter-by-family}: we can make a state-of-the-art algorithm perform much better on a
large number of satisfiable instances, whilst performing worse only rarely on satisfiable instances,
and never on unsatisfiable instances.  Further, these properties are \emph{not} shared by
discrepancy searches (\cref{figure:scatter-dds}) or by random value-ordering with restarts
(\cref{figure:scatter-random}). Finally, we show that this strategy is also effective in an
optimisation setting, producing benefits in two maximum common induced subgraph algorithms.

@@

Our $\FuncSty{propagate}$ routine is similarly close to that of \citet{DBLP:conf/aaai/HoffmannMR17},
differing only in that we carry out the filtering for injectivity
(\siplinerangeref{alldiffstart}{alldiffend}) after every assignment rather than once at the end of
the unit propagation loop. (We make this change because the all-different propagator used is not
idempotent, and we wish to avoid an odd interaction when we introduce nogood filtering.)

@@

We begin by verifying that our implementation of the basic \cref{algorithm:sip} is competitive. In
\cref{figure:others} we show the cumulative number of instances solved over time, using our
implementation (both with the degree heuristic, and with the modifications described in the
remainder of this paper), the Glasgow2 and Glasgow3 algorithms from which \cref{algorithm:sip} is derived
\citep{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}, the PathLAD variation of the LAD
algorithm \citep{DBLP:journals/ai/Solnon10,DBLP:conf/lion/KotthoffMS16}, and VF2
\citep{DBLP:journals/pami/CordellaFSV04}, in each case using the original implementation provided
by the algorithm's authors. For each algorithm, the $y$ value
gives the cumulative number of instances which (individually) can be solved in no more than $x$
milliseconds.  The vertical distance between two lines therefore shows how many more instances can
be solved by one solver than another, if every instance is run separately with the chosen $x$
timeout; the horizontal distance shows how many times longer the per-instance timeout would need to
be to allow the rightmost algorithm to succeed on $y$ out of the 14,621 instances (bearing in mind
that the two sets of $y$ instances solved by the two algorithms could be completely different).

@@

We use the Luby scheme \citep{DBLP:journals/ipl/LubySZ93} to determine when to restart. We count the
number of backtracks (that is, when we reach the end of the main $\KwSty{foreach}$ loop on
\siplineref{countrestarts}) to decide when to restart, rather than counting recursive calls.
Following established wisdom, we multiply each item in the Luby sequence by a magic constant---we
used SMAC \citep{DBLP:conf/lion/HutterHL11} to select a value of 660. (We also tried a geometric
restart scheme. The results were less favourable---it appears that value-ordering heuristics benefit
from a very aggressive restart strategy.)

@@

Two recent maximum common induced subgraph algorithms also make use of backtracking search with
degree as a value-ordering heuristic. The k${\downarrow}$ algorithm
\citep{DBLP:conf/aaai/HoffmannMR17} attempts to solve the problem by first trying to solve the
induced subgraph isomorphism problem, and then if that fails, retries allowing a single unmatched
vertex (and thus using weaker invariants), and so on. Due to its similarity to \cref{algorithm:sip},
we can introduce the same bias and restart strategy, requiring only a small amount of jiggery-pokery
to handle wildcards in nogoods.  Meanwhile, the McSplit${\downarrow}$ algorithm
\citep{DBLP:conf/ijcai/McCreeshPT17} uses a constraint programming style search, but with special
propagators and backtrackable data structures that exploit special properties of the problem. The
unconventional domain store used by McSplit${\downarrow}$ precludes the use of arbitrary unit propagation, and so
when introducing restarts, we cannot propagate using nogoods.  Instead, we can only detect when we
are inside an already-visited branch.  We must therefore use the one watched literal scheme instead.

@@

We would like to thank some anonymous people for their comments. These people have names that take
up approximately this amount of space, which is helpful for us getting the page layout right when
anonymising the paper for review.

>>>>

The subgraph isomorphism problem is to find a copy of a small ``pattern'' graph inside a larger
``target'' graph.  The current single strongest subgraph isomorphism solver uses ``highest degree
first'' as a value-ordering heuristic to direct a constraint programming style search
\citep{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16,DBLP:conf/ijcai/McCreeshPT16}. This
heuristic is much better than branching randomly, but is still far from perfect.  We could try to
introduce tiebreaking mechanisms and similar refinements to strengthen the heuristic. However, given
the extreme cost of making an incorrect branching choice early in search, we believe it is more
fruitful to investigate alternatives to simple backtracking (depth-first) search. This paper
evaluates two such approaches: discrepancy search (which performs poorly), and restarts (which are
extremely beneficial). To explore different parts of the search space when restarting, we introduce
a new value-ordering heuristic, which adds a small amount of randomness to the ``highest degree
first'' heuristic.  The effects of these changes are shown in
\cref{figure:scatter-by-family}: we can make a state-of-the-art algorithm perform much better on a
large number of satisfiable instances, whilst performing worse only rarely on satisfiable instances,
and never on unsatisfiable instances.  Further, these properties are \emph{not} shared by
discrepancy searches (\cref{figure:scatter-dds}) or by random value-ordering with restarts
(\cref{figure:scatter-random}). Finally, we show that this strategy is also effective in an
optimisation setting, producing benefits in two maximum common induced subgraph algorithms.

@@

Our $\FuncSty{propagate}$ routine is also close to that of \citet{DBLP:conf/aaai/HoffmannMR17},
differing only in that we carry out the filtering for injectivity
(\siplinerangeref{alldiffstart}{alldiffend}) after every assignment rather than once at the end of
the unit propagation loop. (We make this change because the all-different propagator used is not
idempotent, and we wish to avoid an odd interaction when we introduce nogood filtering.)

@@

We begin by verifying that our implementation of the basic \cref{algorithm:sip} is competitive. In
\cref{figure:others} we show the cumulative number of instances solved over time, using our
implementation (both with the degree heuristic, and with the modifications described in the
remainder of this paper), the Glasgow2 and Glasgow3 algorithms from which \cref{algorithm:sip} is derived
\citep{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}, the PathLAD variation of the LAD
algorithm \citep{DBLP:journals/ai/Solnon10,DBLP:conf/lion/KotthoffMS16}, and VF2
\citep{DBLP:journals/pami/CordellaFSV04}, in each case using the original implementation provided
by the algorithm's authors. For each algorithm, the $y$ value
gives the cumulative number of instances which (individually) can be solved in no more than $x$
milliseconds.  The vertical distance between two lines therefore shows how many more instances can
be solved by one solver than another, if every instance is run separately with the chosen $x$
timeout; the horizontal distance shows how many times longer the per-instance timeout would need to
be to allow the rightmost algorithm to succeed on $y$ out of the 14,621 instances (bearing in mind
that the two sets of $y$ instances could be different).

@@

We use the Luby scheme \citep{DBLP:journals/ipl/LubySZ93} to determine when to restart. We count the
number of backtracks (that is, when we reach the end of the main $\KwSty{foreach}$ loop on
\siplineref{countrestarts}) to decide when to restart, rather than counting recursive calls.
Following established wisdom, we multiply each item in the Luby sequence by a magic constant---we
used SMAC \citep{DBLP:conf/lion/HutterHL11} to select the value 660. (We also tried a geometric
restart scheme. The results were less favourable---it appears that value-ordering heuristics benefit
from a very aggressive restart strategy.)

@@

Having looked at subgraph isomorphism in detail, we now briefly discuss the more general maximum
common induced subgraph problem.  Two recent algorithms for this problem also make use of
backtracking search with degree as a value-ordering heuristic. The k${\downarrow}$ algorithm
\citep{DBLP:conf/aaai/HoffmannMR17} attempts to solve the problem by first trying to solve the
induced subgraph isomorphism problem, and then if that fails, retries allowing a single unmatched
vertex (and thus using weaker invariants), and so on. Due to its similarity to \cref{algorithm:sip},
we can introduce the same bias and restart strategy, requiring only a small amount of jiggery-pokery
to handle wildcards in nogoods.  Meanwhile, the McSplit${\downarrow}$ algorithm
\citep{DBLP:conf/ijcai/McCreeshPT17} uses a constraint programming style search, but with special
propagators and backtrackable data structures that exploit special properties of the problem. The
unconventional domain store used by McSplit${\downarrow}$ precludes the use of arbitrary unit
propagation, and so when introducing restarts, we cannot propagate using nogoods.  Instead, we can
only detect when we are inside an already-visited branch.  We must therefore use the one watched
literal scheme instead, and we also introduce a basic subsumption scheme to prune redundant clauses.

@@

We would like to thank some anonymous people for their comments.
